---
title: "scE2G prediction properties and QC"
author: "Kayla Brand"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    number_sections: true
---

```{r libraries, include=FALSE}
library(plotly)
library(dplyr)
library(DT)
#library(knitr)
```

```{r import_data, echo=FALSE, include=FALSE}
# stats <- readr::read_tsv(snakemake@input$qc_stats)
# preprint <- readr::read_tsv(snakemake@params$reference_clusters)
# tab_template <- snakemake@params$tab_template
stats <- readr::read_tsv("/oak/stanford/groups/engreitz/Users/kaybrand/scE2G/results/CharacterizationMcGinnis_Dataset7/qc_plots/all_qc_stats.tsv")
preprint <- readr::read_tsv("/oak/stanford/groups/engreitz/Users/kaybrand/scE2G/resources/reference_qc_metrics_sheth_qiu_2024.tsv")
tab_template <- "/oak/stanford/groups/engreitz/Users/kaybrand/scE2G/workflow/scripts/prediction_qc/qc_plot_tab_template.Rmd"
```

```{r filter_data, echo=FALSE}
# Filter out weird clusters
filter_data <- function(dataframe, column, threshold, announce=TRUE, remove=TRUE){
  # announce feature requires a cluster column
  if (announce) {
    removed_identifiers <- unique(dataframe$cluster[dataframe[column] < threshold])
    
    # Only print if there is a problem
    if (length(removed_identifiers) > 0) {
      # Format the identifiers as a comma-separated string
      id_string <- paste(removed_identifiers, collapse=", ")
      
      # Create and print the warning once
      stmt <- sprintf("<span style='color: red'>**WARNING: The following identifier(s) have less than %s in %s:\n%s**</span>", format(threshold, scientific=TRUE), column, id_string)
      cat(stmt, "\n")
    }
  }
  if (remove) {
    return(dataframe[dataframe[column] >= threshold, ])
  } else {
    return(dataframe)
  }
}
## Quietly remove BMMC22_ID2_hi_myeloid_prog
preprint <- filter_data(preprint, "fragments_total", 2e6, announce=FALSE, remove=TRUE)
```

```{r helpers, echo=FALSE}
label_key <- c(fragments_total = "# unique ATAC fragments in cluster", 
               cell_count = "# cells in cluster", 
               umi_count = "# RNA UMIs in cluster",
               frag_per_cell = "Mean unique ATAC fragments per cell", 
               umi_per_cell = "Mean RNA UMIs per cell",
               n_enh_gene_links = "# enhancer-gene links", 
               n_enh_elements = "# unique enhancers",
               mean_genes_per_enh = "Mean # genes per enhancer", 
               mean_enh_per_gene = "Mean # enhancers per gene",
               n_genes_with_enh = "# genes with 1+ enhancer", 
               n_genes_active_promoter = "# genes with accessible promoter", 
               n_genes_not_expressed = "# genes below TPM threshold",
               mean_dist_to_tss = "Mean distance to TSS (bp)", 
               mean_enh_width = "Mean width of enhancer element (bp)",
               mean_fragments = "Mean ATAC fragments per cell",
               mean_umis = "Mean RNA UMIs per cell"
               )

assign_model_color <- function(model_name, reference_color = "#C5CAD7") {
  # Define specific model colors
  predefined_colors <- c(
    "multiome_powerlaw_v3" = "#792374",
    "scATAC_powerlaw_v3" = "#006479",
    "multiome_megamap_v3" = "#9b241c",
    "scATAC_megamap_v3" = "#00488d"
  )
  
  # Create an empty named vector
  model_colors <- setNames(character(2), c("Input data", "Sheth, Qiu"))
  
  # Assign reference color
  model_colors["Sheth, Qiu"] <- reference_color
  
  # Assign predefined colors if present
  for (model in names(predefined_colors)) {
    if (model == model_name) {
      model_colors["Input data"] <- predefined_colors[model]
      return(model_colors)
    }
  }
  
  # If the model is not in the predefined colors list
  if (model_colors["Input data"] == "") {
    model_colors["Input data"] <- colors()[sample(length(colors()), length(1))] #"#c5373d"
    return(model_colors)
  }
}

describe_grid_configuration <- function(yColumns, thresholds) {
  return(unlist(
    lapply(yColumns, function(ycol) {
      lapply(names(thresholds), function(xcol) {
        list(
          x_col = xcol,
          y_col = ycol,
          threshold = thresholds[[xcol]]
        )
      })
    }),
    recursive = FALSE
  ))
}

pretty_name <- function(model) {
  pretty_model_names <- c(
    "multiome_powerlaw_v3" = "scE2G (Multiome, powerlaw)",
    "scATAC_powerlaw_v3" = "scE2G (scATAC-only, powerlaw)",
    "multiome_megamap_v3" = "scE2G (Multiome, average Hi-C)",
    "scATAC_megamap_v3" = "scE2G (scATAC-only, average Hi-C)"
  )
  if (model %in% names(pretty_model_names)) {
    return(pretty_model_names[model])
  } else {
    return(model)
  }
}

# Minimum thresholds for clusters
# REQUIRED for scatter_row functions
thresholds <- c(fragments_total = 2e6,
              cell_count = 100,
              umi_count = 1e6)

# for managing when cell count and umi count are not calculated
no_cell_count = sum(stats$cell_count) < 1
no_umi_count = sum(stats$umi_count) < 1
```

```{r scatter, echo=FALSE}
# Adjustable scatter plot basis of most plots)

threshold_descriptor <- function(column) {
  if (grepl("fragments", column)) {
    return("ATAC")
  } else if (grepl("umi", column)) {
    return("RNA")
  } else if (grepl("cell", column)) {
    return("cells")
  } else {
    return("")
  }
}
create_cluster_plot <- function(data, x_col, y_col, 
                                 category_col = "model_name",
                                 plot_order = c(),
                                 x_threshold = NULL, y_threshold = NULL,
                                 axis_labels = NULL, color_map = NULL,
                                 dot_size = 8, plot_index = 1,
                                 threshold_legend_group = "x_threshold",
                                 log_scale_offset=1,
                                 width=NULL, height=NULL) {
  
  # Create determine plotting order if not specified
  if (length(plot_order) == 0) {
    plot_order <- unique(data[[category_col]])
  }
  
  # Compare models or data sources
  if ("model" %in% category_col) {
    model_source <- "Model:"
  } else if ("source" %in% category_col) {
    model_source <- "Source:"
  } else {
    model_source <- "From:"
  }
  
  # Default axis labels if not provided
  if (is.null(axis_labels)) {
    axis_labels <- c(x_col, y_col)
  }
  
  # Create default color map if not provided
  categories <- unique(data[[category_col]])
  if(is.null(color_map)) {
    if (category_col == "model_name") {
      color_map <- assign_model_colors(categories)
    } else if ((category_col == "source") & ("model_name" %in% colnames(data))) {
      color_map <- assign_source_colors(categories, unique(data$model_name))
    } else {
      default_colors <- RColorBrewer::brewer.pal(max(8, length(categories)), "Set1")
      color_map <- setNames(default_colors[1:length(categories)], categories)
    }
  }
  
  # Calculate x-axis range with buffer to include threshold and all points
  # x is always plotted on a log scale.  Add offset in case of zeros
  x_offset <- paste0(x_col, "_offset")
  data <- data %>% mutate(!!sym(x_offset) := !!sym(x_col) + log_scale_offset)
  # adjusted minimum and maximum
  x_min <- min(c(data[[x_offset]], x_threshold), na.rm = TRUE)
  x_max <- max(c(data[[x_offset]], x_threshold), na.rm = TRUE)
  # buffers and ranges
  buffer <- 0.1 * (log10(x_max) - log10(x_min))
  x_range <- c(10^(log10(x_min) - buffer), 10^(log10(x_max) + buffer))
  
  # For use if y has a threshold, calculate y-axis range to include this threshold
  # Also needed for x-threshold height
  y_min <- min(c(data[[y_col]], y_threshold), na.rm = TRUE)
  y_max <- max(c(data[[y_col]], y_threshold), na.rm = TRUE)
  # Determine if y should be plotted on a linear or log scale
  spans_orders_of_magnitude <- ((log10(y_max+1) - log10(y_min+1)) > 1)
  if (spans_orders_of_magnitude) {
    y_offset <- paste0(y_col, "_offset")
    data <- data %>% mutate(!!sym(y_offset) := !!sym(y_col) + log_scale_offset)
  }
  
  # Determine sizing
  if (is.null(width) & is.null(height)) {
    autosize <- TRUE
  } else {
    autosize <- FALSE
  }
  
  # Plan the y-axis layout
  if (spans_orders_of_magnitude) {
    # Log scale y
    yaxis_layout <- list(
      title = axis_labels[y_col],
      type = "log",
      tickformat = "~s",
      # disables autoscale
      tickvals = 10^(floor(log10(y_min+1)):ceiling(log10(y_max+1)))
    )
  } else {
    # Linear scale
    yaxis_layout <- list(
      title = axis_labels[y_col]
    )
  }
  
  # Create the base plot
  p <- plot_ly(
    width=width,
    height=height
  ) %>%
    layout(
      xaxis = list(
        title = axis_labels[x_col],
        type = "log",
        range = if (is.null(x_threshold)) NULL else log10(x_range),
        #tickvals = 10^(xtick_min:xtick_max),
        tickformat = "~s",
        showgrid = TRUE,
        dtick = 1,
        tick0 = if (!is.null(x_threshold)) floor(log10(x_threshold)) else floor(log10(x_min))
      ),
      yaxis = yaxis_layout,
      hovermode = "closest",
      autosize = autosize
    )
  
  # Add scatter plots for each model
  for (origin in plot_order) {
    model_data <- data %>% filter(!!sym(category_col) == origin)
    
    # Only show legend for the first plot
    show_this_legend <- (plot_index == 1)
    cat("Origin:", origin)
    p <- p %>%
      add_trace(
        data = model_data,
        x = ~get(x_offset),
        y = ~get(if (spans_orders_of_magnitude) y_offset else y_col),
        type = "scatter",
        mode = "markers",
        marker = list(
          color = color_map[origin],
          size = dot_size,
          opacity = if (origin == "Sheth, Qiu") 0.3 else 0.7
        ),
        text = ~cluster,
        hoverinfo = "text+x+y",
        hovertext = ~paste("Cluster:", cluster, 
                           "<br>", x_col, ":", round(get(x_col), 4),
                           "<br>", y_col, ":", round(get(y_col), 4),
                           "<br>", model_source, get(category_col)),
        name = origin,
        legendgroup = origin,  # Group traces with the same name
        showlegend = show_this_legend  # Only show in first plot
      )
  }
  
  # Prevent duplicate threshold legend objects
  if (threshold_legend_group == 'x_threshold') {
    show_threshold_legend <- (plot_index == 1)
  } else {
    show_threshold_legend <- TRUE
  }
  
  # Add threshold line to x-axis if specified
  if (!is.null(x_threshold)) {
    p <- p %>%
      add_trace(
        x = c(x_threshold, x_threshold),
        y = c(
          min(c(data[[y_col]], (if (exists("y_threshold")) y_threshold)), na.rm = TRUE), 
          max(c(data[[y_col]], (if (exists("y_threshold")) y_threshold)), na.rm = TRUE)
          ),
        type = "scatter",
        mode = "lines",
        line = list(color = "black", dash = "dash", width = 1.5),
        name = paste0("Threshold (", threshold_descriptor(x_col), "): ", x_threshold),
        legendgroup = threshold_legend_group,
        hoverinfo = "text",
        hovertext = paste("Threshold:", x_threshold),
        showlegend = show_threshold_legend
      )
  }
  # Add threshold line to y-axis if specified
  if (!is.null(y_threshold)) {
    # Only show threshold in legend for first plot
    show_threshold_legend <- (plot_index == 1)
    
    p <- p %>%
      add_trace(
        #x = c(x_min, x_max),
        x = c(
          min(c(data[[x_col]], (if (exists("x_threshold")) x_threshold)), na.rm = TRUE), 
          max(c(data[[x_col]], (if (exists("x_threshold")) x_threshold)), na.rm = TRUE)
        ),
        y = c(y_threshold, y_threshold),
        type = "scatter",
        mode = "lines",
        line = list(color = "black", dash = "dash", width = 1.5),
        name = paste0("Threshold (", threshold_descriptor(y_col), "): ", y_threshold),
        legendgroup = "y_threshold",
        hoverinfo = "text",
        hovertext = paste("Threshold:", y_threshold),
        showlegend = show_threshold_legend
      )
  }
  
  return(p)
}
```

```{r scatter_grid, echo=FALSE}
# Enables grids with variable y-axes

create_metrics_plot_grid <- function(data, plots_config, axis_labels, nrows, ncols, 
                                     category_col = "model_name",
                                     plot_order = c(), 
                                     thresholds = c(),
                                     color_map = NULL, title = NULL, dot_size = 8,
                                     width=NULL, height=NULL, fontsize=NULL) {
  
  
  # Create default color map if not provided
  categories <- unique(data[[category_col]])
  if(is.null(color_map)) {
    if (category_col == "model_name") {
      color_map <- assign_model_colors(categories)
    } else if ((category_col == "source") & ("model_name" %in% colnames(data))) {
      color_map <- assign_source_colors(categories, unique(data$model_name))
    } else {
      default_colors <- RColorBrewer::brewer.pal(max(8, length(categories)), "Set1")
      color_map <- setNames(default_colors[1:length(categories)], categories)
    }
  }
  
  # Check if configuration is valid
  if(length(plots_config) > nrows * ncols) {
    warning("More plots specified than grid cells available. Some plots will be omitted.")
    plots_config <- plots_config[1:(nrows * ncols)]
  }
  
  # Generate all plots
  plot_list <- list()
  for(i in 1:length(plots_config)) {
    config <- plots_config[[i]]
    
    # Create individual plot, passing the plot index
    plt <- create_cluster_plot(
      data = data,
      x_col = config$x_col,
      y_col = config$y_col,
      category_col=category_col,
      x_threshold = if (config$x_col %in% names(thresholds)) thresholds[[config$x_col]] else NULL,
      y_threshold = if (config$y_col %in% names(thresholds)) thresholds[[config$y_col]] else NULL,
      axis_labels = axis_labels,
      plot_order = plot_order,
      color_map = color_map,
      dot_size = dot_size,
      plot_index = i,
      width=width,
      height=height
    ) %>% layout(
      margin = list(l = 80, r = 50, t = 250, b = 50),
      font = list(size = fontsize)
    )
    
    plot_list[[i]] <- plt
  }
  
  # Create subplot grid with reduced spacing
  subplot_args <- c(
    plot_list,
    list(
      nrows = nrows,
      margin = 0.05,
      titleX = TRUE,
      titleY = TRUE,
      shareX = FALSE,
      shareY = FALSE
    )
  )
  
  grid_plot <- do.call(subplot, subplot_args)
  
  # Configure the layout
  layout_args <- list(
    showlegend = TRUE,
    legend = list(
      orientation = "h",
      xanchor = "center",
      x = 0.5,
      y = -0.1*ncols
    )
  )
  
  # Only add title and padding if title is provided
  if (!is.null(title)) {
    layout_args$title <- list(
      text = title,
      font = list(size = 16),
      y = 0.98
    )
    layout_args$margin <- list(t = 50, b = 80)
  } else {
    layout_args$margin <- list(t = 10, b = 80)
  }
  
  grid_plot <- do.call(layout, c(list(p = grid_plot), layout_args))
  
  return(grid_plot)
}
```

```{r scatter_row, echo=FALSE}
# Scatter plots for row-shaped panels sharing a y-axis

create_cluster_plots_row <- function(data, y_col, category_col = "model_name", plot_order = c(), 
                                     thresholds = NULL, axis_labels = NULL, color_map = NULL,
                                     dot_size = 8) {
  
  # Define the x columns to use for each plot
  input_data <- data %>% filter(source == 'Input data')
  x_columns <- names(thresholds)[sapply(names(thresholds), function(x_col) {
  any(input_data[[x_col]] > 0)
  })] # Will not show plots where the x-axis was not calculated
  
  # Create a list to hold each subplot
  subplot_list <- list()
  
  # Create each subplot
  for (i in seq_along(x_columns)) {
    x_col <- x_columns[i]
    x_threshold <- thresholds[x_col]
    
    # Create custom legend behavior
    plot <- create_cluster_plot(
      data = data, 
      x_col = x_col, 
      y_col = y_col,
      category_col = category_col,
      plot_order = plot_order,
      x_threshold = x_threshold,
      axis_labels = axis_labels,
      color_map = color_map,
      dot_size = dot_size,
      plot_index = i,  
      threshold_legend_group = paste0("threshold_", x_col)
    )
    
    subplot_list[[i]] <- plot
  }
  
  # Combine plots with subplot
  combined_plot <- subplot(subplot_list, nrows = 1, shareY = TRUE, titleX = TRUE)
  
  # Add a unified legend at the bottom in a horizontal orientation
  combined_plot <- combined_plot %>%
    layout(
      legend = list(
        orientation = "h",     # Horizontal legend
        xanchor = "center",    # Center the legend horizontally
        x = 0.5,               # Position at center of plot
        yanchor = "top",       # Anchor at top of the reserved area
        y = -0.25,             # Position below the plots (negative value)
        traceorder = "normal", # Normal trace order
        tracegroupgap = 20,    # Gap between trace groups
        font = list(size = 12), # Font size
        bgcolor = "rgba(255, 255, 255, 0.5)" # Semi-transparent background
      ),
      margin = list(
        b = 120  # Extend bottom margin to make room for the legend
      )
    )
  
  return(combined_plot)
}
```

```{r violin, echo=FALSE}
create_violin_plot <- function(data, x_col, 
                               group_col = "model_name", 
                               axis_labels = NULL,
                               color_map = NULL,
                               point_size = 8,
                               box_width = 0.2,
                               jitter_proportion = 1,
                               violin_width = 0.6,
                               orientation = "h",
                               plot_index = 1,
                               hide_reference_points=FALSE) {
  # Categories to plot
  group_values <- unique(data[[group_col]])
  
  # Default axis labels if not provided
  if(is.null(axis_labels)) {
    axis_labels <- setNames(colnames(data), colnames(data))
  }
  
  # Get the grouping variable and create a default color map if not provided
  if(is.null(color_map)) {
    default_colors <- RColorBrewer::brewer.pal(max(8, length(group_values)), "Set1")
    color_map <- setNames(default_colors[1:length(group_values)], group_values)
  }
  
  # Initialize the plot
  p <- plot_ly()
  
  # Only show legend for the first plot
  show_this_legend <- (plot_index == 1)
  
  # Loop through each group to create violin, boxplot, and points
  for(group_val in group_values) {
    group_data <- data %>% filter(!!sym(group_col) == group_val)
    
    # Control reference data display
    points_setting <- if (group_val == "Sheth, Qiu" && hide_reference_points) "none" else "all"
    jitter_setting <- if (group_val == "Sheth, Qiu" && hide_reference_points) 0 else jitter_proportion
    
    # Add violin plot
    p <- p %>% add_trace(
      data = group_data,
      x = ~get(x_col),
      y = group_val,
      # violins
      type = "violin",
      orientation = orientation,
      width = violin_width,
      fillcolor = adjustcolor(color_map[group_val], 
                              alpha.f = 0.3), #if (group_val == "Sheth, Qiu") 0.3 else 0.4),
      # points
      points = points_setting,
      jitter = jitter_setting,
      pointpos = 0,
      marker = list(color = adjustcolor(color_map[group_val], 
                                        alpha.f = if (group_val == "Sheth, Qiu" && !hide_reference_points) 0.3 else 0.5)),
      # hide the automatic box plot
      box = list(visible = FALSE),
      line = list(color = color_map[group_val], width = 1),
      # metadata
      name = group_val,
      legendgroup = group_val,
      showlegend = show_this_legend,
      text = if ("cluster" %in% colnames(data)) paste("Cluster:", group_data$cluster) else NULL,
      hoverinfo = "text"
    )
    
    # Add boxplot inside the violin
    p <- p %>% add_boxplot(
      data = group_data,
      x = ~get(x_col),
      y = group_val,
      orientation = orientation,
      boxpoints = FALSE,
      boxmean = TRUE,
      width = box_width,
      #color = color_map[group_val],
      line = list(color = "black", width = 2),
      fillcolor = "transparent",
      name = group_val,
      legendgroup = group_val,
      showlegend = FALSE,
      hoverinfo = "none"
    )
    
    # # Add points
    # p <- p %>% add_trace(
    #   y = group_val, #rep(group_val, nrow(group_data)),
    #   x = group_data[[x_col]],
    #   type = "scatter",
    #   mode = "markers",
    #   marker = list(color = color_map[group_val], size = point_size, opacity = 0.6),
    #   name = group_val,
    #   legendgroup = group_val,
    #   showlegend = FALSE,
    #   text = if ("cluster" %in% colnames(data)) paste("Cluster:", group_data$cluster) else NULL,
    #   hovertext = if ("cluster" %in% colnames(data)) 
    #     paste("Value:", round(group_data[[x_col]], 2), "<br>Cluster:", group_data$cluster) 
    #   else paste("Value:", round(group_data[[x_col]], 2)),
    #   hoverinfo = "text"
    # )
  }
  
  # Set the layout
  p <- p %>% layout(
    xaxis = list(
      title = axis_labels[x_col],
      zeroline = FALSE
    ),
    yaxis = list(
      title = "",
      zeroline = FALSE,
      showticklabels = TRUE,
      categoryarray = rev(group_values),
      categoryorder = "array"
    ),
    margin = list(l = 100, r = 50, b = 100, t = 50),
    hovermode = "closest",
    showlegend = show_this_legend,
    legend = list(
      orientation = "h",
      xanchor = "center",
      x = 0.5,
      y = -0.15
    )
  )
  
  return(p)
}
create_violin_plot_grid <- function(data, xVars, group_col, axis_labels, 
                                    nrows=3, 
                                    color_map = NULL, title = NULL,
                                    jitter_proportion = 0.6, violin_width = 1, 
                                    box_width = 0.2, point_size = 6,
                                    hide_reference_points=TRUE) {
  # Create default color map if not provided
  if(is.null(color_map)) {
    color_map <- assign_model_colors(unique(data[[group_col]]))
  }
  
  # Generate all plots
  plot_list <- list()
  for(i in 1:length(xVars)) {
    x_col <- xVars[[i]]
    
    # Create individual plot, passing the plot index
    plt <- create_violin_plot(
      data = data,
      x_col = x_col,
      group_col = group_col,
      axis_labels = axis_labels,
      color_map = color_map,
      #show_legend = FALSE,
      point_size = point_size,
      jitter_proportion = jitter_proportion,
      box_width = box_width,
      violin_width = violin_width,
      plot_index = i,
      hide_reference_points=hide_reference_points
    )
    
    plot_list[[i]] <- plt
  }
  
  # Create subplot grid with reduced spacing
  subplot_args <- c(
    plot_list,
    list(
      nrows = nrows,
      margin = 0.05,
      titleX = TRUE,
      titleY = FALSE,
      shareX = FALSE,
      shareY = TRUE
    )
  )
  
  grid_plot <- do.call(subplot, subplot_args)
  
  # Configure the layout
  layout_args <- list(
    showlegend = TRUE,
    legend = list(
      orientation = "h",
      xanchor = "center",
      x = 0.5,
      y = -0.1
    )
  )
  
  # Only add title and padding if title is provided
  if (!is.null(title)) {
    layout_args$title <- list(
      text = title,
      font = list(size = 16),
      y = 0.98
    )
    layout_args$margin <- list(t = 50, b = 80)
  } else {
    layout_args$margin <- list(t = 10, b = 80)
  }
  
  grid_plot <- do.call(layout, c(list(p = grid_plot), layout_args))
  
  return(grid_plot)
}
```

## Dataset metrics
These plots show the number of ATAC fragments, RNA UMIs, and cells for each cell cluster.

<i>
Note: 

* Any cell types with values below the recommended minimum cluster size for running scE2G are flagged below. Ideally, you will see no red text.
</i>

```{r warnings, echo=FALSE, results='asis'}
## Flag problematic clusters
stats <- filter_data(stats, "fragments_total", 2e6, announce=TRUE, remove=FALSE)
rotten_fruit <- filter_data(stats %>% filter(cell_count>0), "cell_count", 100, announce=TRUE, remove=FALSE)
rotten_fruit <- filter_data(stats %>% filter(umi_count>0), "umi_count", 1e6, announce=TRUE, remove=FALSE)
```


```{r cell_count_warning, results = 'asis', echo=FALSE}
empty_cells <- stats %>% filter(cell_count == 0)
# Check if empty_cells has rows
if (nrow(empty_cells) > 0) {
  # Alert text with Markdown for italics
  alert <- "* *Cell counts are not available for the following clusters:*"
  
  # Create a bullet point list of each empty_cells$cluster, also in italics and indented
  bullet_points <- paste0("  * *", empty_cells$cluster, "*")
  
  # Combine all text into a single string with newline breaks
  output_text <- paste(alert, paste(bullet_points, collapse = "\n"), sep = "\n")
  
  # Output the formatted text
  cat(output_text, "\n")
} else {
  # Display nothing if empty_cells has no rows
  cat("")
}
```


```{r umi_count_warning, results = 'asis', echo=FALSE}
empty_cells <- stats %>% filter(umi_count == 0)
# Check if empty_cells has rows
if (nrow(empty_cells) > 0) {
  # Alert text with Markdown for italics
  alert <- "* *RNA UMI counts are not available for the following clusters:*"
  
  # Create a bullet point list of each empty_cells$cluster, also in italics and indented
  bullet_points <- paste0("  * *", empty_cells$cluster, "*")
  
  # Combine all text into a single string with newline breaks
  output_text <- paste(alert, paste(bullet_points, collapse = "\n"), sep = "\n")
  
  # Output the formatted text
  cat(output_text, "\n")
} else {
  # Display nothing if empty_cells has no rows
  cat("")
}
```


```{r dataset_metrics_set_up, echo=FALSE}
one_model_name <- unique(stats$model_name)[1]

dataset_metrics_df <- 
  stats %>%
    filter(model_name == one_model_name) %>%
    mutate(source = 'Input data') %>%
    mutate(mean_fragments = fragments_total/cell_count, 
         mean_umis = umi_count/cell_count)
```

### Total sequencing depth
```{r data_metrics_ATAC_UMI, echo=FALSE, results='asis', fig.width=4, fig.height=4, out.width="100%"}
if (sum(dataset_metrics_df$umi_count) < 1) {
  # There are no UMI counts
  cat("This plot displays the total ATAC fragments per cluster.  Clusters that fall below the minimum threshold of 2 million ATAC fragments per cluster do not meet the minimum sequencing depth recommended for use in scE2G.")
  create_cluster_plot(dataset_metrics_df,
                    "fragments_total", "umi_count", 
                    plot_order = c("Sheth, Qiu", "Input data"),
                    category_col = "source",
                    x_threshold = 2e6, 
                    y_threshold = NULL,
                    axis_labels = label_key, 
                    color_map = c(
                      "Input data" = "#1C2A43"
                      )
                    )
} else {
  cat("This plot displays total RNA UMIs per cluster by total ATAC fragments per cluster.  Clusters that fall outside the 1st quadrant defined by the minimum threshold of 2 million ATAC fragments and 1 million RNA UMIs do not meet the minimum sequencing depth recommended for use in scE2G.")
  create_cluster_plot(dataset_metrics_df,
                    "fragments_total", "umi_count", 
                    plot_order = c("Sheth, Qiu", "Input data"),
                    category_col = "source",
                    x_threshold = 2e6, 
                    y_threshold = 1e6,
                    axis_labels = label_key, 
                    color_map = c(
                      "Input data" = "#1C2A43"
                      )
                    )
}
```

### Average sequencing depth per cell
```{r dataset_metrics_per_cell, echo=FALSE, results='asis', fig.width=6, fig.height=4, out.width="100%"}
if ((sum(dataset_metrics_df$cell_count) > 0) & (sum(dataset_metrics_df$umi_count) > 0)) {
  cat("These plots display the average count of ATAC fragments per cell and average count of RNA UMIs per cell in each cluster.  It is not recommended to run scE2G on clusters which fall to the left of the 100 cells threshold line.")
  
  dataset_config_alt <- list(
  list(
    x_col = "cell_count",
    y_col = "mean_fragments"
  ),
  list(
    x_col = "cell_count",
    y_col = "mean_umis"
  )
)

# Create the gene metrics grid plot
create_metrics_plot_grid(
  data = dataset_metrics_df,  
  category_col="source",
  plots_config = dataset_config_alt,
  axis_labels = label_key,
  thresholds=c(fragments_total = 2e6,
               cell_count = 100,
               umi_count = 1e6),
  nrows = 1,
  ncols = 3,
  color_map = c(
    "Input data" = "#1C2A43"
  )
)
} else {
  cat("The cell count per cluster was not calculated for these data.  For reference, clusters run in scE2G should have at least 100 cells each.")
}
```

## Prediction properties table
The following table shows aggregate properties of the scE2G predictions for each model and each cell type.  Swipe to see more columns.
```{r interactive_table, echo=FALSE}
# Interactive data table
DT::datatable(stats %>% 
                select(cluster, model_name, everything()) %>%
                mutate(across(where(is.numeric), ~round(., 2))),
              extensions= c("FixedColumns", "Scroller"),
              options = list(
                fixedColumns = list(leftColumns = 3),
                scrollX = TRUE,
                scrollY = TRUE #"400px"
                )
              )
```

```{r create_tabs, echo=FALSE, results='asis'}
# Get unique model names
models <- unique(stats$model_name)

# Generate the tabs content
tabs_text <- paste(
  "## Model prediction property plots {.tabset}\n",
  paste(
    sapply(models, function(m) {
      knitr::knit_expand(file = tab_template, model = m)
    }),
    collapse = "\n\n"
  )
)

# Write the expanded content to a temporary file
tmp_file <- tempfile(fileext = ".Rmd")
writeLines(tabs_text, tmp_file)
```

```{r, child=tmp_file, echo=FALSE}
```

## {-}

This report was generated using [this R Markdown document](https://github.com/EngreitzLab/scE2G/blob/hover_plots/workflow/scripts/prediction_qc/hover_qc_plots.Rmd) and [this template for generating tabs for each model](https://github.com/EngreitzLab/scE2G/blob/hover_plots/workflow/scripts/prediction_qc/qc_plot_tab_template.Rmd).
